### 2018-11-28 paper/paper.tex
adding orcids

\documentclass[modern]{aastex62}

\include{preamble}
\newcommand{\Mdwarf}{Barnard's Star\xspace} % or should it be Barnard's star? or GJ 699?

\setlength{\parindent}{1.4em} % trust in Hogg
\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust in Hogg

\shorttitle{\wobble}
\shortauthors{Bedell et al.}

\graphicspath{ {figures/} }
\DeclareGraphicsExtensions{.pdf,.eps,.png}

\title{\textsc{\wobble: a data-driven method for precision radial velocities}}

\author[0000-0001-9907-7742]{Megan Bedell}
\affiliation{\flatiron}

\author[0000-0003-2866-9403]{David W. Hogg}
\affiliation{\flatiron}
\affiliation{Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\affiliation{Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011, USA}
\affiliation{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg}

\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affiliation{\flatiron}

\author[0000-0001-7516-8308]{Benjamin T. Montet}
\altaffiliation{Sagan Fellow}
\affiliation{\chicago}

\author[0000-0002-0296-3826]{Rodrigo Luger}
\affiliation{\flatiron}

\correspondingauthor{Megan Bedell}
\email{E-mail: mbedell@flatironinstitute.org}

\begin{abstract}
% Context
Extreme-precision radial velocity (\EPRV) exoplanet surveys face considerable data analysis challenges in extracting maximally precise \RVs from spectra.
Chief among these, particularly for the upcoming generation of red-optimized spectrographs targeting M dwarfs, is the presence of telluric absorption features, which are not perfectly known.
Another major limitation on the achievable \RV precision is the need to adopt an imperfect stellar template against which to cross-correlate or otherwise match the observations.
In both cases, the precision-limiting reliance on external information can be sidestepped using the data itself.
% Aims
Here we propose a data-driven method to simultaneously extract precise \RVs and infer the underlying stellar and telluric spectra using a linear model (in the log of flux).
The model employs a convex objective and convex regularization to keep the optimization of the spectral components fast.
We implement this method in \wobble, an open-source \python package which uses \TF in its first non-neural-network application to astronomical data.
In this work, we demonstrate the performance of \wobble on archival \HARPS spectra.
% Results
We recover the canonical exoplanet 51 Pegasi b, detect the secular \RV evolution of the M dwarf \Mdwarf, and retrieve the \RM for the Hot Jupiter HD 189733b.
We also present extremely high-\SNR composite stellar spectra and detailed time-variable telluric spectra derived from these data.
\end{abstract}

\keywords{atmospheric effects, methods: data analysis, planets and satellites: detection, stars: individual (51 Pegasi, Barnard's Star, HD 189733), techniques: radial velocities}

\section{Introduction}

Precise radial velocity (\RV) measurements are critical to the discovery and characterization of exoplanets.
On the order of one dozen dedicated spectrographs exist for the purpose of \RV planet-hunting, with at least as many more currently under construction \citep{Wright2017}.
However, significant challenges exist in deriving precise \RV measurements from these spectra.
%The current capabilities of extreme-precision radial velocity (\EPRV) instruments do not extend to the 10 \cms\ regime,
%In this work, we present an open-source code for \RV\ determination. We use a highly flexible linear model to extract \RV s in a fully data-driven way. %Our method simultaneously models the stellar and telluric spectra.

One of the primary contributors to the noise budget in \RV measurements is the incomplete treatment of telluric features in the
Earth's atmosphere \citep{Halverson2016}.
Often, particular sections of a spectrum that are likely to feature telluric features are identified before the velocity shift of the stellar spectrum is inferred.
These regions are then removed from analysis, leaving only seemingly telluric-free regions to be analyzed \citep[e.g.][]{AngladaEscude2012}.

This method has two significant issues.
The first is that removing sections of the spectrum can remove significant amounts of information about the star, lowering the precision at which we can measure the stellar radial velocity.
Many of the regions of significant telluric absorption lie in the red-optical and near-infrared, where there are abundant narrow spectral features that can be used to improve \RV precision.
This is especially true for M dwarfs, which peak in emitted energy at $\approx 1 \mu$m and have many narrow molecular absorption features in their photospheres \citep{Figueira2016}.
Eliminating large chunks of these spectra will therefore significantly inhibit our ability to detect planets around M dwarfs through \RVs.

Secondly, not all telluric features are obvious.
The Earth's atmosphere induces many small-amplitude features, often referred to as ``microtellurics,'' which are not obvious by eye but can affect the star's inferred \RV at the $\sim 1$ \ms level \citep{Cunha2014}.
As the locations of these features are not known \textit{a priori} and may not even be apparent in stacked spectra of many observations, these spectral regions cannot be thrown out.
Instead, alternative methods to account for these features as a part of the model must be developed and employed in order to mitigate the effect of the Earth's atmosphere on the measured stellar radial velocities.

One such approach is modeling the telluric spectrum using existing line databases like \acronym{HITRAN} \citep{HITRAN2016}.
The telluric model may then be divided out from the observations, assuming the line spread function of the instrument is known \citep[e.g.][]{Seifahrt2010}.
This method relies on existing physical knowledge about the Earth's atmosphere and can be fine-tuned using local observatory measurements of e.g. atmospheric water vapor content \citep{Baker2017}.
However, line databases are incomplete even in significant absorption features when compared to actual observations and certainly do not include microtellurics, making them poorly suited for extreme precision RV applications \citep{Bertaux2014}.

Another option is the use of telluric standard observations: a spectrum of a rapidly rotating early-type star, which is virtually featureless due to extreme rotational line broadening, may be used as a telluric model and divided out.
This approach has the advantage of naturally reproducing the instrumental line profile and current observing conditions if the standard star has a line-of-sight vector sufficiently close to the target and if both observations are taken close together in time.
For these conditions to be true, though, requires a significant investment of observing time, which planet search programs often cannot afford.
Additionally, artifacts may remain near strong telluric features due to the imperfect correction of unresolved features \citep{Bailey2007}.

An alternative approach is the simultaneous modeling of both telluric and spectral features from the data.
As the Earth's motion around the barycenter of the Solar System induces a Doppler shift considerably larger than both the motion of telluric features and the size of a single pixel on the detector, these two spectra can be disentangled.
This process is well-established in the analysis of binary star systems through the development of linear models \citep[e.g.][]{Simon1994} and in a Gaussian process framework \citep{Czekala2017}.
In these cases, both spectra are assumed to be unchanging in time, which is a reasonable approximation of a stellar spectrum but not necessarily of the telluric spectrum.
A more complicated model with time variability in the telluric spectrum may provide a more accurate fit.
Work by \citet{Artigau2014} demonstrates that a Principal Component Analysis (\acronym{PCA}) approach is an effective way of parameterizing telluric spectral variability as a low-dimensional model derived from observational data.
\citet{Artigau2014} use a library of telluric standard observations, which requires a significant investment of observing time to build up, but in principle such a data-driven model should be possible to derive from typical stellar observations if the same star is observed many times at different barycentric \RV shifts.

Just as imperfect telluric modeling can be a noise source in \EPRV analyses, the choice of stellar template can also be a major source of error.
For stabilized, non-gas-cell \RV spectrographs, a standard approach has been to adopt a quasi-binary mask consisting of weighted top-hat functions at the expected locations of informative stellar absorption lines and cross-correlate this mask with the observed spectrum \citep[e.g.][]{Baranne1979, Pepe2002}.
This approach is limited by the accuracy of the mask, and since most masks are built for a broad category of spectral type rather than customized for the individual star in question, it is unlikely that this technique retrieves maximally precise \RVs.
Deriving a custom spectral template by stacking all spectra iteratively as the \RVs are determined has been shown to be a superior approach for stars with complex spectra \citep{AngladaEscude2012}.

Data analysis pipelines for absorption cell instruments have long used a stellar template that is customized to the star in question, although traditionally this template is a single observation taken for the purpose \citep[e.g.][]{Butler1996}.
Higher quality templates can be derived iteratively from the data \citep{Sato2002, Gao2016}.
Such data-driven templates have the benefit of being both customized to the star in question and high-\SNR without requiring observing overhead.
%\todo{(explain how this differs from what we'll do - v similar in spirit although ours is a full forward model with stellar spectrum as a latent variable)}

%\todo{(talk about lack of open-source code and difficulty of modifying \RV pipelines here?)}

Here we develop a data-driven model to infer the telluric and stellar spectra and calculate the stellar \RV at each observed epoch.
The telluric model component may vary with time in a low-dimensional manner, which is also inferred from the data.
Our model requires no prior knowledge of the star or the Earth's atmosphere.
As such, it does not yield absolute measurements of \RVs, only highly precise relative measurements between epochs.

Our approach is similar to that taken by \citet{Gao2016}, but differs in that we take the stellar spectrum as a latent variable to be optimized rather than iteratively determining it by stacking model residuals.
In this sense, ours is a true forward model.

In this work, we focus on the ultra-stabilized spectrograph case, i.e. a reliable instrumental calibration and no absorption cell.
We also assume that multiple epochs of observations exist and that these epochs are spread out across the observing season(s).
This assumption is necessary to enable the disentangling of telluric features from the stellar spectrum.
In this sense our method is intended as a post-processing step, not a real-time data reduction service.
However, the implementation that we present here is designed for flexibility and easy extensibility, and we discuss potential ways to overcome these limitations.

In Section \ref{s:methods}, we outline the model and its key underlying assumptions.
We present an open-source implementation of this method in \python and \TF called \wobble.
In Section \ref{s:results}, we demonstrate \wobble's capabilities by applying our method to \HARPS archival data for three target stars: the planet-hosting solar analog 51 Peg, the quiet M dwarf \Mdwarf, and the Hot Jupiter host HD 189733.
We look further into the detailed time-variable telluric spectra inferred from these data in Section \ref{s:tellurics}.
We revisit many of the assumptions underlying \wobble in Section \ref{s:future} with a detailed look at the limitations of the current implementation.
We outline potential ways of adapting \wobble\ for such cases as instruments with absorption cells, intrinsic time variability in the stellar spectrum, and lower-quality data.
Finally, we conclude with a brief summary in Section \ref{s:conclusion}.

\section{Methods}
\label{s:methods}
\subsection{Model Assumptions}
\label{s:assumptions}

The model underpinning \wobble is designed to be flexible and easily extensible to a variety of situations.
However, a few assumptions are made in this work to simplify the implementation, and we outline those here.
Many of these assumptions could be eliminated with relatively straightforward modifications to the method.
We revisit these in Section \ref{s:future}.

First, we assume that the wavelength calibration and spectral extraction of the instrument are perfect: that is, we begin at the stage of having 1D extracted spectra and corresponding wavelength grids in hand and we do not model any corrections to the wavelength solution.

We assume that the spectra can be modeled as the product of a finite number of components.
For the cases shown in this work, two components are used: a stellar spectrum which is invariant in shape but may be Doppler-shifted, and a telluric spectrum which is fixed to the observatory rest frame but varies in shape.
We choose to work in log(flux) space so that the data to be modeled are simply a sum of the component spectra.

For the case of the telluric component, whose spectrum is allowed to vary with time, we assume this spectral variability is low-dimensional.
This assumption is physically motivated in the sense that that a relatively small number of molecular species contribute to the telluric absorption spectrum.
It is also needed in a practical sense, since every additional dimension over which the telluric spectrum can vary adds several thousand more free parameters to the model.

We assume that the stellar spectrum is invariant with time.
This assumption does not hold true in detail and will be revisited in Section \ref{s:future}.

We assume that both the stellar and telluric spectra are approximately located at zero in logarithmic flux (or unity in linear) with small deviations due to absorption lines.
As a result of this assumption, we are able to apply L1 and L2 regularization to the spectral templates.
Regularization is a commonly used technique in machine learning, where large numbers of free parameters are standard.
It is equivalent to applying a prior to the parameters which pushes them toward zero in the absence of strong evidence otherwise from the data.
The strength of the regularization may be tuned to suit the data at hand through a cross-validation scheme: for example, the best-suited regularization in the bluest spectral orders may be much stronger for the telluric spectrum, where few features are present, than it is for the star, which generally has a dense forest of spectral lines.
The exact implementation and validation of regularization in this model is further described in Section \ref{s:model-eqns}.

Aside from this regularization to push model components to zero in the low-\SNR regime, the model makes no assumptions about the shape of the stellar or telluric spectra.
We solve for spectral templates for each component as a series of control points with no imposed correlations between them, meaning that the line spread function, covariances between lines arising from the same species, and other such physically expected correlations are not built into the model but must be learned in the process of optimizing.
In addition to keeping this model simple, linear, and largely convex, this means that no physical knowledge about the object being observed is needed to extract its \RVs.

For practical applications, the number of epochs $N$ must be large and span a significant fraction of the observing season.
If the number of spectra is small or if all spectra are taken over a very small range of time (e.g. a single night), it will be nearly impossible to disentangle the components of the model, as this process relies on the apparent Doppler shifting of the star(s) and Earth's atmosphere relative to each other.
We will revisit this restriction in Section \ref{s:future}.

Finally, we assume that, when the stellar and telluric components are properly optimized, only Gaussian noise remains.

\subsection{Model Specification}
\label{s:model-eqns}

We take the data to be the $M \times N$ matrix Y, where each entry $y_{m,n}$ is the observed logarithmic flux $\ln f$ for pixel $m$ of $M$ at epoch $n$ of $N$.
We also have a corresponding $M \times N$ matrix of wavelength solutions.
Again, the entries of this matrix are in logarithmic form ($\ln \lambda$).
This choice makes the Doppler shift an additive term, considerably simplifying the math to follow.

We model each data column $y_n$ as the sum of the stellar and telluric spectra:
\begin{equation}
y_n = y_{\star, n} + y_{t, n} + noise.
\end{equation}

The stellar spectrum contribution is:
\begin{equation}
y_{\star, n} = P(v_{obs, n}) \mu_{\star},
\end{equation}
where $P$ is an interpolation operator which applies a Doppler shift by velocity $v_n$ and $\mu_{\star}$ is a spectral template with length $M'$ corresponding to the length of an arbitrarily-chosen wavelength grid of control points.
Because we work in log-wavelength space, the Doppler shift due to velocity $v$ is simply an additive term:
\begin{equation}
 \ln \lambda(v) = \ln \lambda_{0} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}
The $P$ operator adds this term to the $M'$-length wavelength grid for the template spectrum before linearly interpolating the template to the $M$-length data space.

The apparent stellar \RV, $v_{obs, n}$, is a sum of the star's actual velocity about its center of mass ($v_{\star, n}$) and the projected motion of the Earth about the Solar System barycenter (Barycentric Earth Radial Velocity or \BERV), the latter of which is known.

The telluric spectrum contribution is:
\begin{equation}
y_{t, n} =  A_n(\mu_{t} + W_{t} z_n).
\end{equation}
In addition to its mean spectrum $\mu_t$, the telluric component also includes a time-dependent term assembled from two variables: $W_t$, a matrix of ``basis vectors'' for the span of telluric spectral variations, is weighted by $z_n$ to form the spectral contribution at epoch $n$.
$W$ has the shape $M' \times K$ and $z_n$ is a $K$-vector, where $K$ is the number of basis vectors used.
For the purposes of this work we found good performance with $K$ set to 3, but this may vary for other applications.
The net telluric spectrum (mean + time-variable components) is weighted by the airmass at the time of observation $A_n$, a known quantity.

We evaluate a log-likelihood for each epoch
\begin{equation}
\ln \mathcal{L}_n = \sum_{m} -0.5 (y_{m,n} - y'_{m,n})^T C_{m,n}^{-1} (y_{m,n}-y'_{m,n}),
\end{equation}
comparing our model $y'$ to the data $y$, with $C$ representing the covariance matrix of uncertainties for the data.

The number of free parameters in this model is large: we must optimize every grid point in the mean spectral templates and telluric basis vectors along with the stellar \RV and the telluric basis weights for each epoch.
We deal with potential over-fitting issues by applying L1 and L2 regularization to the spectral templates, as discussed in Section \ref{s:assumptions}.

L1 normalization adds a term to the log-likelihood that takes the form \todo{(TO DO: change notation to match literature)}:
$$ L1(p, \lambda) = \lambda \sum_{i} | p_{i} | ,$$
where $p$ is the vector of parameters to be normalized (in this case $\mu_{{\star}}$, $\mu_{{t}}$, or $W_{t}$) and $\lambda$ is the regularization amplitude.

Similarly, L2 normalization adds a term of the form:
$$ L2(p, \lambda) = \lambda \sum_{i} p_{i}^2 .$$

The effectiveness of the regularization depends sensitively on the value of $\lambda$ used: if $\lambda$ is too high, real features will be lost as the parameters are forced to zero, whereas setting $\lambda$ too low will make the regularization ineffective, leaving the model vulnerable to overfitting.

We set regularization amplitudes for \wobble using a cross-validation scheme.
In brief, we set aside 12.5\% of the total epochs as a validation set and, using some value of $\lambda$, run the model optimization on the remaining epochs (the training set).
The resulting best-fit spectral templates and basis vectors are taken as fixed and the time-dependent terms only (\RVs and basis weights) are optimized for the validation epochs.
The $\chi^2$ for the validation epochs can then be adopted as a goodness-of-fit measurement for the $\lambda$ value, and the procedure is repeated for different $\lambda$s to choose the best regularization amplitude.
Because we have multiple regularization terms and multiple amplitudes to set, we begin by hand-setting all amplitudes to a reasonable starting guess and optimize each amplitude sequentially with cross-validation.

With regularization included, the final model likelihood is:
\begin{equation}
\begin{split}
\ln \mathcal{L} = & \sum_{n} \sum_{m} -0.5 (y_{m,n} - y'_{m,n})^T C_{m,n}^{-1} (y_{m,n}-y'_{m,n})  \\
 & + L1(\mu_{\star}, \lambda_{L1, \mu_{\star}}) + L2(\mu_{\star}, \lambda_{L2, \mu_{\star}}) + L1(\mu_{t},  \lambda_{L1, \mu_{t}}) + L2(\mu_{t},  \lambda_{L2, \mu_{t}}) \\
 & + L1(W_t, \lambda_{L1, W_t}) + L2(W_t, \lambda_{L1, W_t}) + L2(z_t, 1.0) ,
\end{split}
\end{equation}
where the regularization amplitude for the basis weights is arbitrarily set to 1.0 and all other regularization amplitudes $\lambda$ are set by grid searches using the above-described validation procedure. \todo{(HOGG: explain why regularization on basis weights is needed)}

%In practice, the most effective regularization amplitudes can vary by orders of magnitude depending on the characteristics of the spectral region being used.
%For instance, both $\lambda_{L1, \mu_{t}}$ and $\lambda_{L2, \mu_{t}}$ need to be much higher when the spectra in question span relatively blue wavelengths with few actual telluric features to fit than they do for the near-infrared wavelength regions with strong telluric features.
%Similarly, the regularization amplitudes for the star tend to be higher for the G dwarf than they are for the M dwarf, which has far more absorption features in its spectra.
%Setting the regularization amplitudes is therefore a crucial first step when applying the \wobble model to data.

\subsection{Optimizing the Model}
\label{s:optimizing}

As an initial guess, we set the star to be stationary, e.g. $v_{obs} = $ \BERV at all epochs.
We may then initialize the stellar spectrum $\mu_{\star}$ by Doppler-shifting the data and calculating the median flux across \BERV-corrected spectra in bins at each model wavelength $\lambda'$.
The telluric template is initialized similarly by using the residuals after the stellar contribution has been removed, again binning in model wavelength (this time without applying any Doppler shift to the spectra) and taking the median values of each bin.
Finally, we initialize the telluric spectrum's basis vectors $W_t$ and weights $z_t$ by performing Principal Component Analysis (\acronym{PCA}) on the residuals after both the stellar spectrum and the mean telluric spectrum have been removed.
The $K$ highest eigenweights and their corresponding eigenvectors are taken as the basis weights and vectors.

Once all parameters are initialized, we optimize them iteratively.
The likelihood function is maximized first by varying the stellar and telluric templates (including the telluric basis vectors).
This step is a convex optimization, meaning that a global optimum for the template parameters should be reached under the condition of fixed time-dependent variables.
Next, the templates are held fixed and the time-dependent parameters (stellar \RVs and telluric basis weights) are varied to maximize the likelihood function again.
\todo{(HOGG: how could this be made convex too?)}
We repeat this procedure, optimizing spectra and time-dependent parameters in turn, until the likelihood appears to converge, typically within 100 iterations.

The inverse variance uncertainty on the parameters is calculated as the Hessian, or the second derivative of the likelihood with respect to each parameter. \todo{(DFM: help me rephrase if needed?)}

\subsection{Combining Spectral Orders}
\label{s:combining-orders}

Most \EPRV instruments are echelle spectrographs spanning many orders.
These orders are often treated as independent spectra when extracting \RVs, and the \RVs for each order are then combined in some manner to get a final time series.

We follow this precedent and optimize the \wobble model individually for each order.
While some \RV precision gain may be possible by inferring a global Doppler shift from all orders simultaneously, it is not clear that the gain would be worth the added computational expense, so we leave this for future work.

After obtaining an $N$-epoch set of observed stellar \RVs from each of the $R$ spectral orders, we combine them by modeling each \RV $v_{n,r}$ as a combination of time-dependent stellar \RV, order-dependent \RV offset, and a Gaussian noise term:
\begin{equation}
v_{n,r} = v_n + v_r + \mathcal{N}(0,\,\sigma_{n,r}^{2} + \delta_r^2 )\,,
\end{equation}
where $v_n$ and $v_r$ are the characteristic \RVs at epoch $n$ and order $r$, $\sigma_{n,r}$ is the estimated measurement uncertainty on $v_{n,r}$, and $\delta_r$ is an additional jitter term specific to order $r$.

\todo{(To do: fill in details: prior on $<v_r>$, optimization over $\delta_r$ with $v_n$ and $v_r$ analytically solved for.)}

\subsection{Implementation}
\label{s:implementation}

The above-described model can be implemented in a variety of ways. We chose to build our code, \wobble, in \code{python} using \TF.
\TF is a \todo{(DFM: write something technical here please)}.
While it has primarily been used in the context of deep learning in the astronomical literature, its auto-differentiation and fast linear algebra capabilities make it valuable for this application as well.
The necessary optimizations over many parameters can be performed with high efficiency by \TF: the below-described analysis of 91 spectra of 51 Peg runs in \todo{30} minutes on a standard Mac desktop.

Our code made open-source under the \acronym{MIT} license and is publicly available on GitHub.\footnote{\url{https://www.github.com/megbedell/wobble}}

\section{Application to \HARPS Data}
\label{s:results}

For the purposes of this work, we chose to test the performance of \wobble on archival spectra from the High Accuracy Radial Velocity Planet Searcher (\HARPS) spectrograph \citep{Mayor2003}.
\HARPS has operated continuously since 2003 and as such has an extensive catalog of publicly available data.
Furthermore, its excellent instrumental stability and careful calibration make the data ideally suited to our method, which relies on having an accurate wavelength solution for every spectrum.

All data were obtained from the \acronym{ESO} public data archive in the form of ``\code{e2ds}'' spectra.
These data are blocks of 72 order by 4096 pixel extracted 1D spectra.
The airmass of the observation and the calculated \BERV are both provided in the \acronym{FITS} header of the spectral data file.
We obtained wavelength solutions for each spectrum from the dedicated \HARPS calibration archive, also maintained by \acronym{ESO}.
\todo{Uncertainty estimates for each pixel were calculated as... [HOGG, help me fill this in - relation to \SNR and conversion to log]}

Before running the above-described model optimization on the data, we first mask out any unreliably measured spectral regions and do a continuum normalization on each spectrum.
This masking is done by setting the inverse variance on the masked data point to zero, effectively removing it from the fit.
In brief, we mask pixels whose extracted flux is below zero as well as regions at the edges of the spectral orders where the local \SNR falls below 5.
After masking these data, we convert the flux to log space and continuum normalize by fitting a polynomial to an asymmetrically clipped subset of the data.

The final stellar \RV retrieved by \wobble is corrected for intra-night drift in the wavelength solution by subtracting off a drift term specified in the data header, consistent with the standard \HARPS pipeline. \todo{(TO DO: make sure this is implemented!)}

Below we describe the results of applying this procedure and optimizing the \wobble model using \HARPS data for stars of different types: a G dwarf with a known planet, an \RV-quiet M dwarf, and an early-K dwarf undergoing a planetary transit event.

\subsection{51 Pegasi}

For \wobble's first test, we chose the first known exoplanet host: 51 Pegasi.
This target is a Sun-like star hosting a planet with an orbital period of 4 days and a mass of 0.5 Jupiter masses \citep{Mayor1995}.
Its canonical status as the first exoplanet discovered means that large amounts of data exist for this system.
In particular, archival \HARPS data exist mainly from efforts to observe reflected-light spectra of the planet \citep{Martins2015}.

We ran \wobble on these archival data to test its performance on recovering a planetary signal with well-known orbital characteristics.\footnote{Based on data obtained from the ESO Science Archive Facility under request numbers mbedell:329968 and mbedell:331551.}
The 91 publicly available spectra in the \HARPS archive are largely concentrated on a few nights of intensive observing, but these nights are sufficiently spread out throughout the year for a wide enough range in \BERV to disentangle the stellar and telluric spectra.
Their \SNR is generally high, ranging from 100 up to 300 pix$^{-1}$ at the central wavelength regions of \HARPS.

Despite the sparse phase coverage available across the planetary orbit, we recover a signal at the correct period and semi-amplitude in the \RVs (Figure \ref{fig:51peg_planet}). \todo{(numbers here, address the structure in residuals)}

Fits to an individual spectrum are shown in Figure \ref{fig:51peg_spectrum}.
We emphasize that no \textit{a priori} information on e.g. expected spectral line positions and depths were used.
The current realization of the \wobble algorithm treats each control point of the template spectra as independent and has no line shape parameterization included.
Nevertheless, the optimized stellar templates clearly reproduce the expected appearance of the spectra in a variety of regimes, from a crowded-line region to a sparser, continuum-dominated region to an extremely strong absorption line like the H$\alpha$ feature.

\begin{figure}
\centering
\includegraphics[width=4in]{51peg_planet}
\caption{Phased orbit of the Hot Jupiter 51 Peg b as recovered in the \wobble\ \RVs. While the phase coverage of \HARPS observations is sparse, we nonetheless recover orbital parameters consistent with those found by other \RV campaigns \todo{(add numbers)}.  \todo{check on: The subtle trends seen near secondary eclipse (phase 0.75) are suggestive of contributions by reflected light from the planetary dayside.}}
\label{fig:51peg_planet}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=5.3in]{51peg_multispectrum}
\caption{Three example echelle orders at a randomly chosen epoch from the \HARPS observations of 51 Pegasi. Data are shown as dark circles, while the \wobble predictions for the stellar and telluric spectral contributions at this epoch are red and blue lines, respectively. Even in the presence of a strong stellar absorption feature such as the H$\alpha$ line (seen at 6561.5 \ang due to the star's Doppler shift at the plotted epoch), small telluric features are clearly recovered. Residuals after subtracting both star and telluric models are shown below each spectrum.}
\label{fig:51peg_spectrum}
\end{figure}

\subsection{\Mdwarf}
\label{s:Mdwarf}

Next, we tested \wobble with 306 epochs of \Mdwarf spectra.\footnote{Based on data obtained from the ESO Science Archive Facility under request numbers mbedell:359445 and mbedell:359501.}
\Mdwarf is a mid-M dwarf, and as such its output is low in the optical; the typical \SNR of the spectra in the central wavelength regions of \HARPS ranges from 20-50.
In practice, this means that the 7 bluest echelle orders were consistently masked and dropped from the fit in accordance with the \SNR criteria outlined previously.
Regardless, a large amount of data remained, allowing us to test \wobble's performance on observations with substantially different properties from the previous case.

\Mdwarf is one of the nearest stars to us at a distance of less than 2 parsecs.
It is also the star with the highest known proper motion.
Its trajectory translates to a projected secular change in \RV of approximately 4.5 $\mathrm{m}~\mathrm{s}^{-1}~\mathrm{yr}^{-1}$ \citep{Kurster2003}.
This trend has been observed in some data sets, including long-term \HARPS and \acronym{HIRES} observations, although \acronym{UVES} \RVs were inconsistent with the predicted linear slope \citep{Kurster2003, Bonfils2013, Choi2013, Montet2014}.
Aside from this linear trend, \Mdwarf is commonly used as an \RV standard M dwarf because little stellar activity has been observed and no planets discovered until very recently despite considerable \RV monitoring \citep{Ribas2018}.

\begin{figure}
\centering
\includegraphics[width=5in]{barnards_rvs}
\caption{Radial velocity measurements for \Mdwarf from \wobble (black) and from the standard \HARPS pipeline (red). The median \RV has been subtracted from each data set. The predicted secular change in \RV due to \Mdwarf's projected motion, calculated using \gaia properties, is shown as a solid blue line (top panel). Residuals away from the predicted trend are plotted in the lower panel. \todo{(TO DO: add error bars and change color scheme)}}
\label{fig:barnards_rvs}
\end{figure}

The \RVs found by \wobble are in excellent agreement with the predicted secular motion (Figure \ref{fig:barnards_rvs}).
Using \gaia parallax and proper motion measurements and following the calculations outlined in \citet{Kurster2003}, we find a secular trend with a slope of 4.53 $\mathrm{m}~\mathrm{s}^{-1}~\mathrm{yr}^{-1}$, deviating from this linearity by less than 1 $\mathrm{cm}~\mathrm{s}^{-1}~\mathrm{yr}^{-2}$ during the decade of \HARPS observations \citep{gaia2016, gaia2018}.

After subtracting the \RVs as predicted by the equations of motion at each epoch, the residuals have low dispersion, as expected for a quiet star with no known planets.
The \acronym{RMS} scatter among \wobble\ \RVs is 2.0 \ms.
This compares favorably to the scatter of 2.5 \ms among \RVs produced by the standard closed-source \HARPS pipeline, although we note that many of the residuals have similar non-zero values in both independently produced RV estimates.
It is possible that these deviations are a real physical effect in the stellar spectrum, but we caution that approximately \ms-level errors are likely being introduced by the barycentric correction.
Currently we assume that the \BERV provided by the native \HARPS pipeline is correct, but this will not hold true if the sky coordinates entered by the observer deviate from the actual on-sky location of the target.
This is especially likely to be the case for high proper motion targets like \Mdwarf.
Indeed, we do note a strong peak at year-long periods in a Lomb-Scargle periodogram of the residual \RVs.
Thus the \RV scatter derived in this work is only an approximate upper limit on the true precision achievable by \wobble. \todo{(TO DO: recalculate \BERV ourselves)}

Testing \wobble on the \Mdwarf data allows us to evaluate its performance in a significantly different regime: the mid-M dwarf spectra are far denser in spectral features than a Sun-like star, and individual observations are at a much lower \SNR than the 51 Peg observations used above, although the total number of spectra is greater.
The resulting spectral fit is shown in Figure \ref{fig:barnards_spectrum}.

\begin{figure}
\centering
\includegraphics[width=5.5in]{barnards_multispectrum}
\caption{Example echelle orders at a randomly chosen epoch from the \HARPS observations of Barnard's Star, plotted as in Figure \ref{fig:51peg_spectrum}.}
\label{fig:barnards_spectrum}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=6in]{barnards_model}
\caption{A subset of the \Mdwarf spectrum including the sodium doublet at 4456-7 \ang. The spectral template derived by \wobble is shown as a red line, while a theoretical \PHOENIX model for an M dwarf with \teff = 3200 K, \logg = 5.0, and \mh = $-0.5$ is shown as a blue line. The \wobble model has been Doppler-shifted to match the rest-frame line positions. \todo{(FIX COLORS!)}}
\label{fig:barnards_model}
\end{figure}

One consequence of working in the low-\SNR regime is that the power to resolve very small telluric features is reduced.
To keep the template spectra from overfitting the noise, the regularization amplitudes on the templates must be raised by several orders of magnitude relative to what was optimal for the 51 Peg fit above, particularly in orders where no strong spectral features are available.
This has the side effect of flattening out weak features if no strong lines are present in a given order.
As a result, we are able to retrieve fewer telluric lines in the bluer orders (compare the top panels of Figures \ref{fig:51peg_spectrum} and \ref{fig:barnards_spectrum} \todo{-- To do: check if there's a better example order}).
In redder orders, where the \SNR is higher and more strong telluric lines are present, the features are retrieved (bottom panel of Figure \ref{fig:barnards_spectrum}).
However, telluric variability was not able to be resolved in the overwhelming majority of orders; in fact, the results presented here were obtained with the tellurics model set to $K=0$ (no variability included).
We therefore caution that difficult-to-resolve microtellurics and time-variable features may not be reliably disentangled from the stellar spectrum when \wobble is applied to lower signal observations \SNR $\lessapprox 50$ pixel$^{-1}$).

Despite the extreme noise at the bluer end of the wavelength range, where \Mdwarf is very faint, \wobble successfully retrieves a template spectrum that appears consistent with general expectations for a mid-M dwarf through most of the \HARPS wavelength range.
As a sanity check, we compare the \wobble stellar template model with a high-resolution \PHOENIX model at \Mdwarf's previously measured spectral parameters \citep[\teff = 3200 K, \logg = 5.0, and \mh = $-0.5$;][and references therein]{Husser2013, Artigau2018}.
Even in the bluest regions, the features inferred by \wobble compare well with the \PHOENIX predictions in broad strokes (Figure \ref{fig:barnards_model}).
\todo{(TO DO: ask Travis Barman about opacity binning affecting the detailed features)}

These results emphasize the potential value of \wobble data products for spectral characterization and model testing, particularly in the case of faint stars and M dwarfs, for which many spectra must be combined to get a reasonably detailed composite spectrum.
The \wobble algorithm is a simple yet robust method for doing such multi-spectra stacking while preserving telluric-contaminated regions and marginalizing over unknown \RV shifts.

\subsection{HD 189733}

\begin{figure}
\centering
\includegraphics[width=5in]{hd189_rm}
\caption{\RVs measured by \wobble for a single night of observations of the Hot Jupiter host HD 189733. The \RM is clearly seen as the planet transits the star. \todo{[blah blah starry blah]}}
\label{fig:hd189_rm}
\end{figure}

While the majority of \EPRV measurements are made for the purposes of observing stellar reflex motion due to planetary orbits, several other applications to exoplanet characterization exist for such high-precision time-series spectra.
One of these uses is measuring the \RM, in which the apparent stellar \RV is observed during planetary transit as a way of mapping the stellar surface and learning the spin-orbit inclination of the system \citep[e.g.][]{Queloz2000, Winn2005}.
While the \RM technically manifests in the stellar spectrum as a distortion to the line profile rather than a true Doppler shift, it is typically detected using similar methods to a standard \RV analysis.

We apply \wobble to a data set consisting of a single transit of the hot Jupiter HD 189733b to test its sensitivity to stellar line asymmetries as well as its performance in the regime of a single night's observations.
The observations in question consist of a consecutive series of 40 spectra, each with an \SNR $\sim 90$, taken on the night of August 28, 2007.\footnote{Based on data obtained from the ESO Science Archive Facility under request numbers \todo{x}.}

Because all observations came from a single night, there will be no significant shift of the stellar spectrum with respect to the telluric spectrum and the power of \wobble to disentangle the two is severely limited.
For this reason, we fixed the telluric spectrum to a constant template.
We derived this template by running \wobble on the 51 Peg data with non-time-variable tellurics and adopting the resulting high-quality time-invariant telluric spectrum.
The stellar spectrum of HD 189733 and its \RVs were left as free parameters.

The resulting \RV signal retrieved by \wobble is shown in Figure \ref{fig:hd189_rm}.
We fit the signal using \starry \citep{Luger2018}. \todo{(RODRIGO: fill in a bit about methods)}
\todo{(comparison of retrieved parameters to previous work here)}

These results confirm that the \wobble method measures line asymmetries as \RV shifts in the same manner as traditional \RV analysis techniques.
While this is a useful approximation in the case of the \RM, in general the confusion of these two spectral changes is a major cause of correlated noise in \RV time series \citep[e.g.][]{}.
The simplicity and extensibility of the \wobble method is useful in this regard, as the model could be modified to fit line asymmetries and true Doppler shifts as separate parameters.
This would help to disentangle signals caused by photospheric features like starspots from the \RV signal due to reflex motion in a planetary system.
We discuss this prospect further in Section \ref{s:future}.

\section{Telluric Features}
\label{s:tellurics}

\begin{figure}
\centering
\includegraphics[width=6.5in]{telluric_comparison}
\caption{Fits to a 15-\ang region with substantial telluric contamination for 51 Peg (left) and \Mdwarf (right). The upper panels show the \wobble best-fit telluric spectrum (blue line), best-fit stellar spectrum (red line), data (black points), and the sum of the telluric and stellar model predictions (black line). Residuals are shown in the lower panels. Although the fits to each data set were performed entirely independently, the optimized telluric spectra are nearly identical, demonstrating that \wobble successfully finds an accurate representation of the common telluric absorption spectrum. \todo{(TO DO: add y-label, fix centering)}}
\label{fig:telluric_comparison}
\end{figure}

In the above analyses, telluric spectra were inferred independently for the 51 Peg and \Mdwarf data sets.
However, in principle the telluric features should be common to all \HARPS spectra, an assumption that we used for the case of HD 189733.

We show a comparison between 51 Peg and \Mdwarf telluric fits in Figure \ref{fig:telluric_comparison}.
Although they were fit using entirely different data sets with dissimilar stellar spectra, the resulting telluric fits are indeed extremely similar.
This is generally true for all wavelength regions in which tellurics are detected.
The only regime in which the comparison fails is for low-\SNR spectral orders of the \Mdwarf data, where the telluric spectra are featureless due to the strong regularization, as discussed in Section \ref{s:Mdwarf}.

The physically motivated expectation that all observations should share a common telluric component could be built into the \wobble model.
For the fit to HD 189733, for example, we fixed the telluric spectrum to use the model inferred from the better-studied star 51 Peg.
A more robust, albeit more computationally expensive, approach would be to fit many stars simultaneously so that their shared telluric spectrum can be inferred using all the available data.
Given enough spectra, this approach should yield an incredibly detailed model of telluric features and their time variability.
We leave such an effort to future work.

\begin{figure}
\centering
\includegraphics[width=5in]{telluric_standard}
\caption{Comparison of telluric spectrum derived by \wobble analysis of 51 Peg spectra (upper panel) and the spectrum observed with a telluric standard star (lower panel) around an H$_2$O absorption band. No telluric standard star data was used in determining the \wobble telluric template. The \wobble method delivers a very high \SNR telluric spectrum with no overhead observing time required.}
\label{fig:telluric_standard}
\end{figure}

As another test of our derived telluric spectra, we compared the mean template from 51 Peg to an observation of the telluric standard star \todo{HR 3090 (+details)}.\footnote{Based on data obtained from the ESO Science Archive Facility under request numbers mbedell:385782 and \todo{xxx}.}
As shown in Figure \ref{fig:telluric_standard}, the spectra agree well.
Moreover, the composite telluric spectrum inferred from the time series of 51 Peg observations is at a much higher \SNR than the single shot telluric standard spectrum, so that low-amplitude lines which border on statistical insignificance in the standard star show up clearly in the \wobble results.
Unlike a traditional telluric standard observation, no overhead time is required to produce these results.

Of course, one strength of a telluric standard star is that its spectrum can capture the true telluric spectrum at any given moment in time, while the telluric template determined by \wobble is time-averaged and may not perfectly capture the telluric absorption lines in any given observation.
For this reason, we included time-variable components in the model.
A physical interpretation of these components should be possible.

\begin{figure}
\centering
\includegraphics[width=5in]{telluric_basis2}
\caption{Example of the mean telluric spectral template in a small wavelength region (upper panel) and the three basis vectors used to capture the time-variable contributions to the spectrum, staggered by an arbitrary offset for display (lower panel).}
\label{fig:telluric_basis}
\end{figure}

An example of the inferred variability from the 51 Peg analysis is shown in Figure \ref{fig:telluric_basis}.
There is clearly coherent structure in the variations.
As we might expect, many of the lines co-vary with each other: as all known telluric lines in this region arise from H$_2$O, this likely corresponds to time-variable levels of atmospheric water vapor content.
In general, the basis vectors are suggestive of Doppler shifts, perhaps due to wind speeds in different layers of the atmosphere contributing individually to the overall line shapes.
These hypotheses could be tested in the future by seeking correlations between basis weights and externally-measured weather diagnostics.
%\todo{(is this enough? also show that basis weights change seasonally-ish?)}

\section{Limitations and Extensibility of \wobble}
\label{s:future}

In this work we have tested \wobble exclusively using \HARPS data, which allowed us to make relatively strict assumptions about the data quality.
We have also made a number of restrictive assumptions about the underlying physical model, for example taking each astronomical source to be a single star.
However, there are many other \EPRV data sets which violate some of these assumptions but could nevertheless benefit from the analysis techniques used by \wobble.

We now turn our discussion to ways in which other data sets may violate the assumptions made in this work and the corresponding modifications to \wobble that would enable it to operate more effectively in these regimes.
For simplicity, we break this discussion down into two categories: changes to the quality or type of spectroscopic data used and changes to the model.
We additionally outline some general changes or enhancements that could be made to potentially improve the \RV precision achieved by \wobble.

\subsection{Changes to the Data}
\label{s:data-changes}

In general, the data used in this work are high quality in terms of \SNR, spectral resolution, and quantity.
As demonstrated for the case of \Mdwarf, \wobble is limited in its operability when the data are at low \SNR.
\wobble implicitly assumes Gaussian noise in logarithmic flux, which is not strictly accurate.
This may become a substantial issue at low \SNR, making \wobble a poor tool choice for accurate modeling of very low \SNR spectra.
\todo{(HOGG: maybe say something about how we could mitigate this with changes to continuum handling?)}

Our method may also fail at low spectral resolution.
In particular, if the stellar spectrum does not shift with respect to the telluric spectrum by at least a resolution element, it will become extremely challenging to disentangle the two components.
\todo{(HOGG: say other things about undersampled lines?)}

Along with minimum requirements on the spectral data quality, \wobble has some limitations driven by the number of spectra available and the sampling of those spectra.
Some minimum quantity of spectra are necessarily simply because the spectral templates contribute many free parameters, deriving up the data points needed for inference.
The sampling of those spectra are also important because the stellar lines must undergo substantial \RV variations to disentangle them from the telluric features, which is only feasibly achieved by observing throughout the season to take advantage of the $\sim 30$ \kms \BERV shift.
This requirement could be partially mitigated by modeling multiple stars with a shared telluric spectral template, which we discuss in Section \ref{s:improvements}.
However, in general the \wobble method fundamentally requires a substantial number of high quality spectra to perform reliably.

Another key assumption made in this work that there is no gas cell represented in the data.
While this is true for \HARPS, other instruments like \acronym{HIRES}, \acronym{APF}, and \acronym{PFS} include a gas cell set in optical path to imprint its absorption spectrum on the observed spectra \todo{(cite)}.
Applying \wobble to spectra that include the imprint of a gas cell is technically trivial.
The gas absorption lines are fixed features at the observatory rest frame which multiply into the spectrum.
This makes them nearly indistinguishable from the mean telluric spectrum, and they should in fact be absorbed into this component.
If the observations in question cover a range of different airmasses, it may be necessary to explicitly add a third model component to represent the gas cell; this component would be identical to the tellurics without the airmass scaling.

Of course, absorption cells in general exist because the instantaneous calibration of the spectrograph in question is not reliable at the required \RV precision level.
This means that a gas cell instrument will generally violate other, more critical assumptions about the data quality.

It is likely that for such an instrument, our assumption that the wavelength solution is perfect does not hold.
One simple way to deal with this would be to allow the observatory rest frame component(s) of the model to change in \RV.
The stellar \RV measurement would then be differential with respect to the effective spectrograph \RV.
This effective spectrograph \RV may vary in different spectral orders or smaller wavelength regions. \todo{(HOGG: say a thing)}
\todo{(TO DO: add some citations to using tellurics for this)}

%Spectral variability due to LSF changing? Additional parameters to the P operators, this would go into the non-convex optimization part (ask Hogg to throw in some words earlier on about location vs linear parameters)
Another way a less well-calibrated instrument may violate \wobble's assumptions is in the non-negligible variability of all spectral lines due to changes in the instrumental line spread function (\LSF).
In principle, if the changes to the \LSF were well-understood they could be hard-coded into the P operators.
Realistically speaking, if \LSF variations are present they are probably not sufficiently well-understood to take this approach; instead, we might want to fit these variations as a data-driven model component.
This leads us into the territory of making more fundamental changes to the \wobble model.

\subsection{Changes to the Model}
\label{s:model-changes}

The test data used in this work were all instances of a single, bright star, so that only one stellar component was necessary to model the astronomical source.
Other data sets may require a more complex model, including additional stars, planetary spectra, or non-negligible sky background.
They may also require more complex treatment of spectral variability.
There are some trade-offs associated with adding these components to the model, which we discuss here.

One important feature of \wobble is that it models all fluxes in the log-space.
When the model consists solely of multiplicative components, as is the case for a single star + telluric absorption, this choice makes the optimization convex at fixed \RV and computationally tractable \todo{(HOGG: edit plz)}.
If we chose instead to convert the $y$ data vectors to linear fluxes and use a model consisting only of additive components, the same equations and algorithms would apply.
Example applications of such a model might be a multiple star system or a faint star with substantial sky background present in the observed spectra.
If, however, the model needed to incorporate both additive and multiplicative components, major structural changes would be necessary and much of the optimization speed would be lost.
This means that at present, additive components may only be easily modeled by either fixing or neglecting the multiplicative components.
\todo{(HOGG: say things about hacks - maybe in a footnote even?)}

One extra model component that is at present both simpler to implement and perhaps more important to the results than additive components is the spectral continuum.
We have assumed that our simple continuum normalization is good, but this is certainly not true at a detailed level.
To some extent, this wrongness is absorbed into the telluric spectrum, as both the tellurics and the continuum can be treated as observatory rest frame features.
However, since the continuum and tellurics vary through time in different ways, the continuum should in principle be its own independent model component.
This component will be highly degenerate with the tellurics, but a clever representation could circumvent this: for example, the continuum spectrum could be placed on an extremely sparse wavelength grid to ensure that very localized effects like absorption lines are not included.

Another way in which the \wobble model may fail for some regimes is in its treatment of spectral variability.

We have assumed that the telluric spectrum is variable in a low-dimensional sense only.
While this assumption has a basis in the physical understanding of telluric lines, which are produced by only a small number of molecular species, it is not guaranteed that this reasoning translates into low variability in the current parameterization.
We may instead wish to change the spectral representation so that additional physics is included, as discussed further in Section \ref{s:improvements}.
Alternatively, we could keep the current representation but fit the telluric model to the residuals away from some physics-based model.

We have also assumed that the stellar spectrum is constant, which is almost certainly false.
A variety of physical effects cause changes in the spectrum including stellar oscillations, convection, and activity in the form of starspots or plages.
\wobble can model stellar variability trivially by making the stellar component of the model take a similar form to the tellurics, where a set of basis vectors and weights are inferred from the data to capture variability.
This \PCA-like approach may be quite effective \citep{Davis2017}.
It does, however, introduce potential degeneracies between the inferred stellar variability and the Doppler shift.
To avoid this degeneracy, we might instead represent the stellar spectrum with an explicit parameterization of the line profile and restrict variability to the line depths, widths, and skews; such a change would fall along the same lines as the global \LSF parameterization proposed in Section \ref{s:data-changes}.

%Some absorption lines are more sensitive to the effects of stellar activity than others (Dumusque). If these fall into a small number of categories, a PCA-like dimensionality reduction should do well. Address pros and cons of introducing Gaussian processes to parameterize line profiles. deal with this similarly to the LSF changing? also, maybe we need to regularize (or put priors on) the \RVs themselves and force them to sum of Keplerians for the purpose of breaking degeneracies with stellar variability.

\todo{(HOGG: Bayesian version of \wobble?)}

\subsection{Basic Adjustments}
\label{s:improvements}

Even working within \wobble's current assumptions and running on \HARPS data, there are potential changes to be made to the implementation of \wobble that may improve its performance.
Many of these are quite basic and straightforward changes to be made within the existing framework of \wobble.
We discuss some of potential adjustments here.

Perhaps the most obvious shortcoming of \wobble is its reliance on regularization.
We chose this framework because it keeps the model convex and its implicit assumptions (that the spectra are a flat continuum by default) closely mirror our physical understanding.
However, setting the regularization amplitudes optimally is by far the computationally slowest and most unwieldy part of applying \wobble to a new data set.
In principle, as data sets get larger, regularization becomes less necessary.
One way to remove the need for telluric regularization might be to run on a data set that is composed of many different stars, each with their own individual spectra and \RVs but with a shared telluric component, so that the data informing the telluric lines is sufficiently large.
This could be done with relatively trivial changes to the \wobble code, although it would require strong computational resources.

One choice that \wobble makes differently from other data-driven \RV codes is the way we represent the underlying spectral components.
We use an extremely simple representation: a model grid of wavelengths and corresponding fluxes that gets linearly interpolated by the P operator.
One possible improvement would be to go to a higher-order interpolation scheme.
Another approach would be a non-parameteric method like a Gaussian process (as in the \code{STARFISH} code from \citet{Czekala2017}).
A third approach would be to use a model that incorporates more physics: this could range from a model with fixed lines that may vary only in terms of common line profile parameters to a complex model including various atomic and molecular physics of the lines.
A sensible way to implement this within the existing \wobble framework might be to add a physically motivated mean spectral template and optimize $y_{\star, n}$ vectors of residuals away from this template.
In any case, the general approach of \wobble does not depend intrinsically on the spectral representation, and swapping out the current representation for something else is simple, provided that the representative function can be implemented in \TF.

Another choice we made which is almost certainly suboptimal is the treatment of each spectral order in the echelle data as an independent spectrum with its own stellar \RV.
In reality, every \RV should be informed by all spectral orders simultaneously.
Implementing this would take on some extra computational cost, but being in \TF will help to mitigate this.

Finally, a change that is quite major but worth stating regardless is the possibility of fitting the data directly in the 2D domain.
Widely used methods of extracting 1D spectra almost certainly sacrifice some degree of spectral information \citep{Bolton2010}.
In the future, we may wish to move to fitting the data directly without extracting.
The equations for \wobble would change very little in this regime: essentially, the P operator would become a 2D interpolator.
However, implementation of this model would become much more complicated, and extra variable components like sky background and spectrograph dispersion functions would need to be introduced.
While this implementation is far beyond the scope of current work, we nevertheless emphasize that the \wobble model is a useful starting framework in which to work towards this goal.

\vspace{2cm}

\section{Conclusion}
\label{s:conclusion}

Extremely precise \RV measurements are a critical part of modern exoplanet detection and characterization.
Many dedicated \EPRV instruments exist and more are being built with the goal of achieving better than \ms precision over long timescales.
Beyond the hardware challenges of building such instruments, software challenges exist as well: extracting maximally precise \RVs from high-quality spectra is not a trivial task.
Moreover, most existing \EPRV pipelines are closed-source, proprietary, and built for use with a specific instrument in mind.
While traditional \RV extraction methods are demonstrably suboptimal when it comes to treatment of telluric features or use of a fixed stellar template spectrum, experimenting with changes to the existing pipelines is generally not possible.

In this work, we have proposed a simple linear model for simultaneously inferring stellar spectra, telluric spectra, and \RVs from the data.
Our model does not rely on physical knowledge of the star or the Earth's atmosphere.
We implement this method in \wobble, an open-source code designed to be extensible and adaptable for a variety of data sets.

By running the \wobble algorithm on archival \HARPS spectra for a variety of stars, we have demonstrated its basic capabilities.
The \RVs achieved through this method are comparably precise to those obtained with the closed-source \HARPS pipeline.
Furthermore, the stellar and telluric spectra inferred purely from the data are high-quality and accurate.
We have shown that consistent telluric features are found from independent data sets with extremely different properties.

The data-driven methods used in \wobble are highly promising for the treatment of microtelluric lines and other deviations from spectral models.
This makes the approach particularly valuable for upcoming infrared surveys targeting M dwarfs, where telluric features are non-negligible and stellar models often fall short of matching the data.
It also eliminates costs in observational overhead for obtaining telluric standard spectra or accurate stellar templates.

We make \wobble freely available to the \RV community to be adapted for use with various instruments and data sets.
We highlight the versatility of the algorithm in Section \ref{s:future} with suggested changes and improvements for applications to different regimes.
It is our hope that openly shared development of next-generation \RV analysis techniques will enable the community as a whole to make full use of the information-rich data being produced by current and future \EPRV surveys.

\software{\todo{TO DO! --} astropy, numpy, matplotlib, scipy, tensorflow}

\acknowledgements{It is a pleasure to acknowledge Jacob L. Bean, Heather Knutson, Timothy Morton, Melissa Ness, Andreas Seifahrt, and Julian St\"{u}rmer for helpful discussions.
We are grateful to Hans-Walter Rix and the Max-Planck-Institut f\"ur Astronomie for their hospitality while developing key parts of this project.

This research has made use of the services of the ESO Science Archive Facility.

This work has made use of data from the European Space Agency (ESA) mission
{\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions
participating in the {\it Gaia} Multilateral Agreement.}

\bibliographystyle{apj}
\bibliography{paper.bib}%general,myref,inprep}

\end{document}